---
title: "p8105_hw2_gw2383"
author: "Guojing Wu"
date: "2018/10/1"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

***

## Problem 0

The folder contains the "entrance and exit for each subway station in NYC" is `Subway_data`.

## Problem 1

### Read, clean, retain and convert the data

```{r}
subway_df = 
  read_csv("./Subway_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% # read data
  janitor::clean_names() %>% # cleaning the name
  select(line:entry, vending, ada, ada_notes) %>% # filter
  mutate(.data = .,
    entry = ifelse(entry == "YES", TRUE, FALSE)) # convert

subway_df$entry[1:100]
```

### About this dataset

The variables this dataset contains are: `r colnames(subway_df)`.

The cleaning step includes reading: `read_csv`, cleanning variable names: `janitor::clean_names`, filter: `select` and convert: `mutate`.

The dimension are `r dim(subway_df)[1]` x `r dim(subway_df)[2]`.

This data is not tidy because it contains variable name "route", all the route names shall be integrated into a single column.

