---
title: "p8105_hw2_gw2383"
author: "Guojing Wu"
date: "2018/10/1"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)
```

***

## Problem 0

The folder contains the "entrance and exit for each subway station in NYC" is `local_data`.

## Problem 1

### Read, clean, retain and convert the data

```{r}
subway_df = 
  read_csv("./local_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% # read data
  janitor::clean_names() %>% # cleaning the name
  select(line:entry, vending, ada, ada_notes) %>% # filter
  mutate(.data = .,
    entry = ifelse(entry == "YES", TRUE, FALSE)) # convert

subway_df$entry[1:100]
```

### About this dataset

The variables this dataset contains are: `r colnames(subway_df)`.

The cleaning step includes reading: `read_csv`, cleanning variable names: `janitor::clean_names`, filter: `select` and convert: `mutate`.

The dimension are `r dim(subway_df)[1]` x `r dim(subway_df)[2]`.

This data is not tidy because it contains variable name "route", all the route names shall be integrated into a single column.

### Some questions

* The number of distinct stations are:
```{r} 
dim(distinct(subway_df, line, station_name))[1]
```

* The number of stations that are ADA compliant is:
```{r}
sum(distinct(subway_df, line, station_name, ada)$ada == TRUE)
```

* The proportion of station entrances / exits without vending allow entrance is:
```{r}
entry_no_vending = subway_df$entry[which(subway_df$vending == "NO")] # station entrances / exits without vending
sum(entry_no_vending == TRUE) / length(entry_no_vending)
```

### Reformat the data

```{r}
subway_tidy_df = 
  gather(subway_df, key = route_number, value = route_name, route1:route11) %>% # gather route by route_number
  filter(!is.na(route_name)) # remove the one without route_number
```

The number of distinct stations serve the A train is `r dim(distinct(subway_tidy_df, line, station_name, route_name) %>% filter(route_name == "A"))[1]`. 

Of the stations that serve the A train, `r dim(distinct(subway_tidy_df, line, station_name, route_name, ada) %>% filter(route_name == "A") %>% filter(ada == TRUE))[1]` stations are ADA compliant.

## Problem 2

### Read and clean the Mr. Trash Wheel sheet

```{r}
trash_df = 
  readxl::read_excel("./local_data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 1, range = "A2:N338") %>% # read
  janitor::clean_names() %>%
  rename(weight_in_tons = weight_tons, volume_in_cubic_yards = volume_cubic_yards) %>% # clean names
  filter(!is.na(dumpster)) %>% # omit rows
  mutate(sports_balls = as.integer(round(sports_balls))) # convert
```

### Read and clean the precipitation data

```{r}
# read precipitation data for 2016
prcp_2016 = readxl::read_excel("./local_data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 5, range = "A2:B14") %>% 
  mutate(Year = 2016) %>% 
  rename(Total_inch = Total) %>% 
  select(Year, Month, Total_inch) %>% 
  filter(!is.na(Total_inch))
# read precipitation data for 2017
prcp_2017 = readxl::read_excel("./local_data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 4, range = "A2:B14") %>% 
  mutate(Year = 2017) %>% 
  rename(Total_inch = Total) %>% 
  select(Year, Month, Total_inch) %>% 
  filter(!is.na(Total_inch))

# combine the dataset
prcp_df = 
  full_join(prcp_2016, prcp_2017, by = c("Year", "Month", "Total_inch")) %>% 
  mutate(Month = month.name[Month])
```

### About these data

* The final version of Mr. Trash Wheel data is not a tidy data, because the `r colnames(trash_df)[2:4]` is redundant.

* The Mr. Trash Wheel data has `r dim(trash_df)[1]` observations and `r dim(trash_df)[2]` variables. The important variables include `dumpster`, `weight` , `volume` and `homes powered`: 

```{r, echo = FALSE}
head(trash_df, 10)[, c(1, 5, 6, 14)]
```

* The precipitation data has `r dim(prcp_df)[1]` observations. And all the variables in this is important:

```{r, echo = FALSE}
head(prcp_df, 10)
```

* The total precipitation in 2017 is `r sum(prcp_2017$Total_inch)`.

* The median number of sports balls in a dumpster in 2016 is `r median(trash_df$sports_balls[which(trash_df$year == 2016)])`.

## Problem 3

### create the dataset

```{r}
brfss_df = 
  p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names() %>% # cleaning
  filter(topic == "Overall Health") %>% # focus on the “Overall Health” topic 
  select(-class, -topic, -question, -sample_size, -confidence_limit_low:-geo_location) %>% # remove
  rename(location_abbr = locationabbr, location_desc = locationdesc) %>% # format the variable name
  spread(key = response, value = data_value) %>% # spread the data by response
  janitor::clean_names() %>%
  select(year, location_abbr, location_desc, poor, fair, good, very_good, excellent) %>% # clean the name again
  mutate(above_average = very_good + excellent) # create a new column
```

### Answer questions

* The number of unique locations included in the dataset is `r dim(distinct(brfss_df, location_desc))[1]`. Calculating the number of distinct states in this dataset: `r dim(distinct(brfss_df, location_abbr))[1]`, we can see that every state is represented. The state that beed observed the most is `r names((brfss_df$location_abbr %>% table() %>% sort(decreasing = T))[1])` with `r (brfss_df$location_abbr %>% table() %>% sort(decreasing = T))[[1]]` observations.

* The median of the “Excellent” response value is `r median((brfss_df %>% filter(year == 2002) %>% select(excellent))$excellent, na.rm = T)`. (The NAs are removed)

* Histogram of “Excellent” response values in the year 2002.

```{r}
ggplot(
  brfss_df %>% filter(year == 2002), 
  aes(x = excellent)
  ) +
  geom_histogram() +
  labs(
    x = "excellent value",
    title = "histogram of excellent value"
  ) +
  theme_bw()
```

* Scatterplot

```{r}
ggplot(
  brfss_df %>% filter(year >= 2002 & year <= 2010) %>% filter(location_desc %in% c("NY - New York County", "NY - Queens County")),
  aes(x = year, y = excellent, color = location_desc)
) +
  geom_point() +
  labs(
    y = "excellent value", 
    title = "proportion of Excellent values in New York County and Queens County"
  ) +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```